---
title: "Homework 1 - Time Series Toolbox"
author: "Tizian Dagner"
date: 'Submission date: 17.12.2020'
output:
  html_document:
    theme: cerulean
    css: styles/homework.css
    toc: yes
    toc_depth: 3
    toc_float: yes
  word_document:
    toc: yes
    toc_depth: '3'
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
options(digits = 5)

# libraries to load
library(fpp3)
library(fpp2)
```

# Exercise 1

For the following series, find an appropriate Box-Cox (?BoxCox) transformation in order to stabilize the variance. First plot the time series and decide which of them need a transformation. 

  * `usnetelec`
  * `usgdp`
  * `mcopper`
  * `enplanements`

```{r}
autoplot(usnetelec)
autoplot(usgdp)
autoplot(mcopper)
autoplot(enplanements)
```


# Exercise 2

Why is a Box-Cox transformation unhelpful for the `cangas` data?

```{r}
autoplot(cangas)
```
Your answer:
Hilfreich bei steigender Varianz über die Zeit, hier jedoch nur temporär höhere Varianz, davor und danach jedoch klein.


# Exercise 3

What Box-Cox transformation would you select for **your** retail data from Exercise 4 in Visualization?

```{r}
# Example 
retaildata <- readxl::read_excel("data/retail.xlsx", skip=1)
myts <- ts(retaildata[,"A3349640L"], frequency=12, start=c(1982,4))
autoplot(myts)
myts %>% BoxCox(lambda = 0) %>% autoplot()
```

From visual inspection, a log transformation would be appropriate here. It also makes sense, as retail expenditure is likely to increase proportionally to population, and therefore the seasonal fluctuations are likely to be proportional to the level of the series. It has the added advantage of being easier to explain than some other transformations. Finally, it is relatively close to the automatically selected value of `BoxCox.lambda(myts)` $= `r round(BoxCox.lambda(myts),3)`$.

If you have selected a different series from the retail data set, you might choose a different transformation.

Your answer:
OPEN------------------------------------------------------------------

# Exercise 4

Calculate the residuals (?residuals) from a seasonal naive forecast applied to the quarterly Australian beer production data from 1992. The following code will help.

```{r}
(?residuals)
help("ausbeer")
beer_forec = snaive(ausbeer, h=40)
residu = residuals(beer_forec)

autoplot(ausbeer)
autoplot(beer_forec)
```

Test if the residuals are white noise and normally distributed.

```{r}
checkresiduals(beer_forec)
```

What do you conclude?

Your answer: Residuals sind nicht nur white noise (Correlogram) -> kein gutes Modell für eine Vorhersage


# Exercise 5

Are the following statements true or false? Explain your answer.

> a. Good forecast methods should have normally distributed residuals.

Your answer: True, einfacher die Vorhersage-Intervalle zu berechnen

> b. A model with small residuals will give good forecasts.

Your answer: True, Residuals = Abweichung Vorhersage zu wahren Werten -> je kleiner die Residuals, desto genauer die Vorhersage

> c. The best measure of forecast accuracy is MAPE.

Your answer: False, abhängig vom Mittleren Abweichungsquadrat

> d. If your model doesn't forecast well, you should make it more complicated.

Your answer: False, eventuell ist das Modell nur durch Ausreißer oder schlechte Vorverarbeitung schlecht.

> e. Always choose the model with the best forecast accuracy as measured on the test set.

Your answer: Nicht eindeutig lösbar, da abhängig von der Ähnlichkeit zwischen Test- und Realdaten

# Exercise 6

For your retail time series (from Exercise 4):

Split the data into two parts using

```{r}
retaildata <- readxl::read_excel("data/retail.xlsx", skip=1)
myts <- ts(retaildata[,"A3349640L"], frequency=12, start=c(1982,4))
myts.train <- window(myts, end=c(2010,12))
myts.test <- window(myts, start=2011)
```

Check that your data have been split appropriately by producing the following plot.

```{r}
autoplot(myts) +
  autolayer(myts.train, series="Training") +
  autolayer(myts.test, series="Test")
```

Calculate forecasts using `snaive` applied to `myts.train`.

```{r}
fc_train = snaive(myts.train)
fc_test = snaive(myts.test)
```

Compare the accuracy of your forecasts against the actual values stored in `myts.test`.
(?accuracy) 

```{r}
(?accuracy)
accuracy(fc_train)
accuracy(fc_test)
```

The number to look at here is the test set RMSE of 71.443. That provides a benchmark for comparison when we try other models.

Check the residuals. Do the residuals appear to be uncorrelated and normally distributed?

```{r}
checkresiduals(fc_train)
```

Your answer: OPEN-------------------------------


How sensitive are the accuracy measures to the training/test split?
OPEN-------------------------------

# Exercise 7

`visnights` contains quarterly visitor nights (in millions) from 1998-2015 for eight regions of Australia.

Use `window()` to create three training sets for `visnights[,"QLDMetro"],` omitting the last 1, 2 and 3 years; call these train1, train2, and train3, respectively.

```{r}
train1 <- window(visnights[, "QLDMetro"], end = c(2015, 3))
train2 <- window(visnights[, "QLDMetro"], end = c(2015, 2))
train3 <- window(visnights[, "QLDMetro"], end = c(2015, 1))
```

Compute one year of forecasts for each training set using the `snaive()` method. Call these `fc1`, `fc2` and `fc3`, respectively.

```{r}
fc1 = snaive(train1)
fc2 = snaive(train2)
fc3 = snaive(train3)
```

Use `accuracy()` to compare the MAPE over the three test sets. Comment on these.
First we will copy the actual data into a variable. Then we can do an accuracy comparison.

```{r}
qld <- visnights[, "QLDMetro"]
accuracy(fc1, qld)
accuracy(fc2, qld)
accuracy(fc3, qld)
```

This should give similar results to this consolidated results table.

```
                     ME     RMSE      MAE        MPE     MAPE      MASE       ACF1
Training set  0.1637836 1.742687 1.360271  0.4384347 7.357322 1.0000000 0.06643175
Test set fc1 -1.3010774 1.301077 1.301077 -6.9956861 6.995686 0.9564839         NA
Test set fc2 0.08383478 1.387447 1.384912 -0.4063445 6.589342 1.019346 -0.50000000
Test set fc3 0.06202858 1.132896 0.9294135 -0.237857 4.425934 0.6738562 -0.51548610
```

The lower MAPE value for "fc3" indicates a better result when we use the previous 3 values for the `snaive()` prediction.

# Exercise 8

Use the Dow Jones index (data set `dowjones`) to do the following:

Produce a time plot of the series.

```{r}
help(dowjones)
# todo
autoplot(dowjones) 
```

Produce forecasts using the drift method and plot them.

Let's assume we want to forecast the next 5, 10 and 15 values.

```{r}
dowfc1 <- rwf(dowjones, drift=TRUE, h=5)
# todo
dowfc2 <- rwf(dowjones, drift=TRUE, h=10)
dowfc3 <- rwf(dowjones, drift=TRUE, h=15)
```

Then we can plot these values.

```{r eval = FALSE}
# remove eval=FALSE
autoplot(dowjones) +
  autolayer(dowfc1, PI=FALSE, series="Drift 5") +
  autolayer(dowfc2, PI=FALSE, series="Drift 10") +
  autolayer(dowfc3, PI=FALSE, series="Drift 15") +
  xlab("Time") + ylab("Closing Price (US$)") +
  ggtitle("Dow Jones index") +
  guides(colour=guide_legend(title="Forecast"))
```

We show that the forecasts are identical to extending the line drawn between the first and last observations.

We can plot the forecasts in a different order, so the shorter forecasts are superimposed, showing the lines are the same.

```{r eval = FALSE}
# remove eval=FALSE
autoplot(dowjones) +
  autolayer(dowfc3, PI=FALSE, series="Drift 15") +
  autolayer(dowfc2, PI=FALSE, series="Drift 10") +
  autolayer(dowfc1, PI=FALSE, series="Drift 5") +
  xlab("Time") + ylab("Closing Price (US$)") +
  ggtitle("Dow Jones index") +
  guides(colour=guide_legend(title="Forecast"))
```

Try using some of the other benchmark functions to forecast the same data set. Which do you think is best? Why?

The time series isn't seasonal, so the seasonal naive method is not viable. However, we can use the mean and naive methods.

```{r}
# todo
fc_mean <- meanf(dowjones)
fc_naive <- naive(dowjones)

autoplot(fc_mean)
autoplot(fc_naive)

accuracy(fc_mean)
accuracy(fc_naive)
```

The three values will be very different here. The Mean will use the data set, so is unlikely to follow the current trendline. -> Naive ist deutlich genauer, da eine deutlich geringere Streuung der Vorhersage vorliegt.

# Exercise 9

Consider the daily closing IBM stock prices (data set `ibmclose`).

Produce some plots of the data in order to become familiar with it.

```{r}
autoplot(ibmclose)
ggAcf(ibmclose)
gglagplot(ibmclose)
```

Split the data into a training set of 300 observations and a test set of 69 observations.

```{r}
ibm.train <- window(ibmclose, end=300)
ibm.test <- window(ibmclose, start=301)
```

Try using various benchmark methods to forecast the training set and compare the results on the test set. Which method did best?

```{r eval = FALSE}
# remove eval=FALSE
h <- length(ibm.test)
m.f <- meanf(ibm.train, h=h)
# todo
rw.f <- naive(ibm.train, h=h)
rwd.f <- rwf(ibm.train, drift=TRUE, h=h) 


autoplot(ibmclose) +
  xlab("Day") +
  ggtitle("Daily closing IBM stock prices") +
  autolayer(m.f$mean, col=2, series="Mean method") +
  autolayer(rw.f$mean, col=3, series="Naive method") +
  autolayer(rwd.f$mean, col=4, series="Drift method")

# accuracy(m.f,ibm.test)
# todo
accuracy(m.f, ibm.test)
accuracy(rw.f, ibm.test)
accuracy(rwd.f, ibm.test) 
```


Check the residuals of your preferred method. Do they resemble white noise?

```{r}
checkresiduals(rwd.f)
```
Weißt Ähnlichkeiten zu white noise auf, jedoch nicht eindeutig identifizierbar

# Exercise 10

Consider the sales of new one-family houses in the USA, Jan 1973 -- Nov 1995 (data set `hsales`).

Produce some plots of the data in order to become familiar with it.

```{r}
print(hsales)
autoplot(hsales)
ggAcf(hsales)
```

Split the `hsales` data set into a training set and a test set, where the test set is the last two years of data.

```{r}
sales.train <- window(hsales, end=1993)
sales.test <- window(hsales, start=1994)

autoplot(sales.train)
autoplot(sales.test)
```

Try using various benchmark methods to forecast the training set and compare the results on the test set. Which method did best?

```{r}
h <- length(sales.test)
salesMean.f <- meanf(sales.train, h=h)
salesNaive.f <- naive(sales.train, h=h)
salesSnaive.f <- snaive(sales.train, h=h)
salesDrift.f <- rwf(sales.train, drift=TRUE, h=h)

accuracy(salesMean.f, sales.test)
accuracy(salesNaive.f, sales.test)
accuracy(salesSnaive.f, sales.test)
accuracy(salesDrift.f, sales.test)
```

In terms of accuracy measures on the test set, the seasonal naive method does better.

Check the residuals of your preferred method. Do they resemble white noise?

```{r eval = FALSE}
# remove eval=FALSE
checkresiduals(salesSnaive.f)
```
Kein eindeutiges Indiz für white noise, Korrelation ändert sich von positiv zu negativ

