---
title: "Homework 1 - Time Series Visualization"
author: "Prof. Dr. Stephan Trahasch"
date: 'Submission date: 17.12.2020'
output:
  html_document:
    theme: cerulean
    css: styles/homework.css
    toc: yes
    toc_depth: 3
    toc_float: yes
  word_document:
    toc: yes
    toc_depth: '3'
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
options(digits = 5)

# libraries to load
library(fpp3)
library(fpp2) 
```

# Exercise 1
Use the help menu to explore what the series `gold`, `woolyrnq` and `gas` represent. These are available in the `forecast` package.

```{r, echo=TRUE, cache=TRUE, eval=FALSE}
help(gold)
# to do
help("gold") #Daily morning gold prices in US dollars. 1 January 1985 - 31 March 1989.
help("woolyrnq") #Quarterly production of woollen yarn in Australia: tonnes. Mar 1965 - Sep 1994.
help("gas") #Australian monthly gas production: 1956-1995.

```

Use `autoplot` to plot each of these in separate plots.

```{r, echo=TRUE, cache=TRUE}
# to do
autoplot(gold)
autoplot(woolyrnq)
autoplot(gas)
```

**What is the frequency of each commodity series?**
Hint: apply the `frequency()` function.

```{r, echo=TRUE, cache=TRUE}
# to do
frequency(gold) #täglich
frequency(woolyrnq) #quartalsweise
frequency(gas) #monatlich
```

Use `which.max()` to spot the outlier in the `gold` series. 
**Which observation was it?**

```{r, echo=TRUE, cache=TRUE}
# to do
which.max(gold) #outlier ist 770
```

# Exercise 2 

Import the file `tute1.csv` from the folder data and review its contents. You should find four columns of information. Columns B through D each contain a quarterly series, labelled Sales, AdBudget and GDP. Sales contains the quarterly sales for a small company over the period 1981-2005. AdBudget is the advertising budget and GDP is the gross domestic product. All series have been adjusted for inflation.

You can read the data into R with the following script:

```{r, echo=TRUE, cache=TRUE}
tute1 <- read.csv("data/tute1.csv", header = TRUE)
View(tute1)
```

We convert the data to time series. The `[,-1]` removes the first column which contains the quarters as we don't need them now.

```{r, echo=TRUE, cache=TRUE}
mytimeseries <- ts(tute1[, -1], start = 1981, frequency = 4)
```

Construct time series plots of each of the three series and use 'facets = TRUE' .

```{r, echo=TRUE, cache=TRUE}
# to do
autoplot(mytimeseries, facets = TRUE)
```

Check what happens when you don't include `facets=TRUE`.

```{r, echo=TRUE, cache=TRUE, dependson="exmyts"}
# to do
autoplot(mytimeseries) #Zeitreihen werden in einem Chart visualisiert
```

# Exercise 3

Import some monthly Australian retail data from the foler data. These represent retail sales in various categories for different Australian states, and are stored in a MS-Excel file.

You can read the data into R with the following script:

```{r, echo=TRUE, cache=TRUE}
retaildata <- readxl::read_excel("data/retail.xlsx", skip = 1)
view(retaildata)
```

The second argument (`skip=1`) is required because the Excel sheet has two header rows.

Select one of the time series as follows (but replace the column name with your own chosen column):

```{r, echo=TRUE, cache=TRUE}
myts <- ts(retaildata[, "A3349640L"], frequency = 12, start = c(1982, 4))
```

Explore your chosen retail time series using the following functions: `autoplot`, `ggseasonplot`, `ggsubseriesplot`, `gglagplot`, `ggAcf`

```{r, echo=TRUE, cache=TRUE}
# to do
autoplot(myts)
ggseasonplot(myts)
ggsubseriesplot(myts)
gglagplot(myts)
ggAcf(myts)
```

**Can you spot any seasonality, cyclicity and trend? What do you learn about the series?**

Your answer:
ANSEWR MISSING -----------------------------------------------------------------------------------------------------------------------------------------------------------------------

# Exercise 4
Create time plots of the following time series: goog, writing, fancy, a10, h02.

    + Use `help()` to find out about the data in each series.
    + For the `goog` plot, modify the axis labels and title.

```{r, echo=TRUE, message=FALSE, warning=FALSE}
# to do
help("goog") #closing stock prices of GOOG
help("writing") #industry sales for printing an writing paper
help("fancy") #monthly sales for a souvenir shop
help("a10") #monthly government expenditure as part of the Pharmaceutical Benefit Scheme for products falling under ATC code A10
help("h02") #monthly government expenditure as part of the Pharmaceutical Benefit Scheme for products falling under ATC code H02

autoplot(goog) +
ggtitle("Closing stock prices of GOOG") +
xlab("Day") + ylab("Price in $")

```

Use the `ggseasonplot` and `ggsubseriesplot` functions to explore the seasonal patterns in the following time series:  `writing`, `fancy`, `a10`, `h02`.

- What can you say about the seasonal patterns?
- Can you identify any unusual years?


```{r, echo=TRUE}
# writing
# to do
ggseasonplot(writing)
ggsubseriesplot(writing)
```

Your answer:Noticeable drop in August


```{r, echo=TRUE}
# fancy
# to do
ggseasonplot(fancy)
ggsubseriesplot(fancy)
```

Your answer:Noticeable increase from October to December


```{r, echo=TRUE}
# a10
# to do
ggseasonplot(a10)
ggsubseriesplot(a10)
```

Your answer: Decrease from January to February, 1991 just has data till June


```{r, echo=TRUE}
# h02
# to do
ggseasonplot(h02)
ggsubseriesplot(h02)
```

Your answer: Extreme decrease from January to February, then increase till December, 1991 just has data from July to December


# Exercise 5
Use the following graphics functions: `autoplot`, `ggseasonplot`, `ggsubseriesplot`, `gglagplot`, `ggAcf` and explore features from the following time series: `hsales`, `usdeaths`, `bricksq`, `sunspotarea`, `gasoline`.

    + Can you spot any seasonality, cyclicity and trend?
    + What do you learn about the series?

Example for hsales:

```{r, echo=TRUE, cache=TRUE}
autoplot(hsales)
ggseasonplot(hsales)
ggsubseriesplot(hsales)
gglagplot(hsales)
ggAcf(hsales)
```

  + Seasonality evident in all plots
  + Cyclicity seen in first two plots
  + No trend
  + ACF only shows seasonality. Cycle length too long to show up here.

```{r, echo=TRUE, cache=TRUE}
# to do
autoplot(usdeaths)
ggseasonplot(usdeaths)
ggsubseriesplot(usdeaths)
gglagplot(usdeaths)
ggAcf(usdeaths)
```

Your answer:


```{r, echo=TRUE, cache=TRUE}
# to do
autoplot(bricksq)
ggseasonplot(bricksq)
ggsubseriesplot(bricksq)
gglagplot(bricksq)
ggAcf(bricksq)
```

Your answer:


```{r, echo=TRUE, cache=TRUE}
# to do
autoplot(sunspotarea)
ggseasonplot(sunspotarea)
ggsubseriesplot(sunspotarea)
gglagplot(sunspotarea)
ggAcf(sunspotarea)
```

Your answer:


```{r, echo=TRUE, cache=TRUE}
# to do
autoplot(gasoline)
ggseasonplot(gasoline)
ggsubseriesplot(gasoline)
gglagplot(gasoline)
ggAcf(gasoline)
```

Your answer:


# Exercise 6

The `pigs` data shows the monthly total number of pigs slaughtered in Victoria, Australia, from Jan 1980 to Aug 1995. Use `mypigs <- window(pigs, start=1990)` to select the data starting from 1990. Use `autoplot` and `ggAcf` for `mypigs` series and compare these to white noise plots.

```{r, echo=TRUE, cache=TRUE}
# White noise
set.seed(1711)
y <- ts(rnorm(50))
autoplot(y) + ggtitle("White noise")
```

```{r}
ggAcf(y)
```


```{r, echo=TRUE, cache=TRUE}
# todo
mypigs <- window(pigs, start=1990)
autoplot(mypigs)
ggAcf(mypigs)
```

Your answer:


# Exercise 7

`dj` contains 292 consecutive trading days of the Dow Jones Index. Use `ddj <- diff(dj)` to compute the daily changes in the index. Plot `ddj` and its ACF. Do the changes in the Dow Jones Index look like white noise?

```{r}
# todo
ddj<-diff(dj)
autoplot(ddj)
ggAcf(ddj)
```

Your answer:

